Analysis of pudding following Davidson 1970

### 2021-11-04 inst/pudding.Rmd
more work

```{r,davidsonvshyper2,cache=TRUE}
ph <- pudding_maxp # ph == 'pudding hankin'; saves typing
pd <- davidson     # pd == 'pudding davidson'
print(ph)
print(pd)
th <- ph[7]        # th == 'tie hankin'
td <- pd[7]        # td == 'tie davidson'
hankin <- matrix(NA,15,3)
davidson <- matrix(NA,15,3)
m <- allbinom(6,2)
m

colnames(hankin) <- c("i","j","tie")
colnames(davidson) <- c("i","j","tie")
rownames(hankin) <- apply(m,2,paste,collapse="")
rownames(davidson) <- apply(m,2,paste,collapse="")
for(n in seq_len(15)){
      i <- m[1,n]
      j <- m[2,n]
      dh <- ph[i] + ph[j] + th    # dh == 'denominator hankin'
      hankin[n,1] <- ph[i]/dh
      hankin[n,2] <- ph[j]/dh
      hankin[n,3] <- th   /dh

      dd <- pd[i] + pd[j] + td*sqrt(pd[i]*pd[j])  # dd == 'denominator davidson'
      davidson[n,1] <- pd[i]/dd
      davidson[n,2] <- pd[j]/dd
      davidson[n,3] <- td*sqrt(pd[i]*pd[j])/dd
}
```

```{r}
print(hankin)
print(davidson)
par(pty="s")
plot(c(t(hankin)),c(t(davidson)),col=c('red','green','blue'),pch=16,asp=1,xlim=c(0.25,0.45),ylim=c(0.25,0.45),type='b')

abline(0,1)
```

>>>>

```{r,davidsonvshyper2,cache=TRUE}
ph <- pudding_maxp # ph == 'pudding hankin'; saves typing
pd <- davidson     # pd == 'pudding davidson'
print(ph)
print(pd)

`hankin` <- function(ph){
  ##  ph a vector of length n+1; first n entries are strengths, last one
  ##  propensity to tie
  n <- length(ph)-1
  m <- allbinom(n,2)
  out <- matrix(NA,nrow=ncol(m),ncol=3)
  colnames(out) <- c("i","j","tie")
  rownames(out) <- apply(m,2,paste,collapse="")
  th <- ph[n+1]
  for(o in seq_len(ncol(m))){
      i <- m[1,o]
      j <- m[2,o]
      dh <- ph[i] + ph[j] + th    # dh == 'denominator hankin'
      out[o,1] <- ph[i]/dh
      out[o,2] <- ph[j]/dh
      out[o,3] <- th/dh
  }
  return(out)
}

`davidson` <- function(pd){

  ##  pd a vector of length n+1; first n entries are strengths, last one
  ##  nu
  n <- length(pd)-1  # how many puddings
  nu <- pd[n+1]  
  m <- allbinom(n,2)
  out <- matrix(NA,nrow=ncol(m),ncol=3)
  colnames(out) <- c("i","j","tie")

  rownames(out) <- apply(m,2,paste,collapse="")

  nu <- pd[n+1]

  for(o in seq_len(ncol(m))){
      i <- m[1,o]
      j <- m[2,o]
      dd <- pd[i] + pd[j] + nu*sqrt(pd[i]*pd[j])  # dd == 'denominator davidson'
      out[o,1] <- pd[i]/dd
      out[o,2] <- pd[j]/dd
      out[o,3] <- nu*sqrt(pd[i]*pd[j])/dd
  }
  return(out)
}

plotter <- function(ph,pd,...){
  h <- hankin(ph)
  d <- davidson(pd)
  nh <- nrow(h)
  plot(c(h),c(d),pch=16,asp=1,col=c(rep("red",nh),rep("blue",nh),rep("green",nh)),...)
}
```

```{r}
print(ph)
print(hankin(ph))
print(pd)
print(davidson(pd))
jj <- c(2,5)/10
par(pty='s')
plotter(ph,pd,xlim=jj,ylim=jj)
grid()
```

Now calculate likelihoods:

```{r}
n <- pudding_table[,-(1:3),]
h <- hankin(ph)
d <- davidson(pd)
n
h
d
sum(n*log(h))
sum(n*log(d))
sum(n*log(h))-sum(n*log(d))
```
